%!TEX root = paper.tex
 \section{Introduction}
%basic idea of message passing with different send modes. The problem of sends modes, its importance, and why it is difficult: message non-determinism.
Message Passing Interface (MPI) is a prevalent programming model in high performance computing (HPC).
The idea of message passing in MPI programs is simple at the outset; A message is communicated by matching a send and a receive in the runtime. Despite the simplicity of the programming model, MPI supports four send modes in message passing that makes the semantics much more complicated.
\begin{compactitem}
\item The regular send does not return until the send buffer can be used. 
\item The buffered send has to buffer the message in the system and returns immediately as it is issued. 
\item The synchronous send does not return until the matching receive is issued in the runtime. 
\item The ready send may returns correctly if the matching receive is already issued in the runtime; otherwise, an error message is sent out.
\end{compactitem}
A problem in any MPI program is the violation of send modes.
In particular, the synchronous send may be incompatible with an MPI program where the send can not return in any possible schedule, therefore makes the program deadlock.  
Also, the ready send may return before the matching receive is issued, where an error message is sent out in the execution. 
%The regular send and buffered send may also be violated due to improper implementation of system buffer. For example, the regular send may return without freeing the send buffer. The buffered send may incorrectly buffer the message. This paper assumes the correct usage of buffering system and focuses on the problem of synchronous send and ready send.
The problem of verifying synchronous send and ready send is essential to any MPI application since it is difficult for a developer to predict. This problem is also much difficult because of the message non-determinism, where concurrent sends to a process may arrive in any order. As such, a receive may potentially match with more than one send in the runtime system. 

%prior work on verifying message passing programs, not able to verify send modes,
Prior work on verifying MPI programs can be roughly grouped into dynamic analysis where an existing runtime is manipulated to explore different scheduling outcomes \cite{DBLP:conf/ppopp/VakkalankaSGK08,DBLP:conf/sbmf/SharmaGB12}, model checking where a model of the original program is analyzed \cite{DBLP:conf/vmcai/Siegel07,DBLP:conf/pvm/Siegel07}, runtime verification where the program execution is observed but not manipulated \cite{DBLP:conf/sc/VetterS00,DBLP:conf/parco/KrammerBMR03,DBLP:conf/ptw/HilbrichSSM09}, and symbolic model checking where a model of the program is symbolically analyzed with an SMT solver \cite{DBLP:conf/kbse/HuangMM13,HuangNFM15}. While many of the solutions are able to verify message non-determinism, None of them is specific to verifying MPI sends modes.


%our solution: SMT
%novelty and the relation to the old algorithm 
This paper looks specifically at symbolic model checking and ways to better leverage the SMT solver in verifying send modes in MPI programs. 
The most efficient types of encoding for symbolic model checking of MPI programs, where efficient means they scale to programs with a high degree of message non-determinism, rely on the concept of a \emph{match pair} and a \emph{concurrent trace program} abstraction (CTP) \cite{DBLP:conf/kbse/HuangMM13,HuangNFM15,DBLP:conf/fm/ForejtKNS14,DBLP:conf/atva/ElwakilYW10,DBLP:conf/issta/ElwakilY10}. 
A CTP is an abstraction of the program that removes all branching so each process is a sequence of send and receive operations \cite{DBLP:conf/kbse/HuangMM13,HuangNFM15}. 
The abstraction is constructed by observing the execution of the program in the runtime for a given input. 
This paper presents a symbolic model checking that gives the SMT solver an encoded CTP with a set of match pairs, and the solver then tries to resolve the encoding to prove the compatibility of synchronous send or to detect the violation of ready send. 

In particular, this paper presents how to encode the behavior of synchronous send and ready send into SMT formulas. The formula size is quadratic. 
As MPI point-to-point communication is also essential to verifying MPI send modes, this paper adapts the existing encoding rules for MPI semantics \cite{HuangNFM15}. 
The new encoding, if satisfiable, is sufficient to prove a CTP to be synchronous send compatible. To detect the violation of ready send, one more step is needed, where the intended erroneous  behavior of ready sends is encoded with the satisfiable encoding. If the combined encoding is also satisfiable, a violation of ready send is detected. To our knowledge, the new encoding in this paper is the first verification approach for solving the problem of send modes.

%why need small match set, 
Another important factor for an efficient SMT encoding is the size of the input match pair set. A large input set typically means an expensive solver call especially if there is a high-degree of message non-determinism. 
The idea in this paper is to reduce the cost of the solver call by reducing the size of the input match set to each call. Such a reduction cannot, however, be done naively; any match pair set must be \emph{message complete}.  A message complete match pair set for a CTP is one that lets the program run to completion (assuming it is free of deadlock). 

%our solution for match generation
This paper present a new algorithm that generates the match pairs in two steps. 
First, each process is sectioned, where a section contains a sequence of receives in the process and a number of sends from all the senders that may potentially match these receives. The number of sends from each sender is roughly bounded by  $k$ given as a user input. 
%For each section, the algorithm precisely adds a set of sequences of sends from multiple potential sender processes. 
The sends from each sender are distributed sequentially. 
The total number of the sends distributed to the section is equal to the number of receives in the same section. 
%The number of the receives depends on three values that are statically known: the positive integer $k$ configured by the user, the number of all the potential sender processes and the number of the unmatched receives in the process. 
Match pairs are then generated for each section independently. This independent sectioning effectively ignores combinations of match pairs, that are feasible, but only available when the considered concurrent outstanding sends are more than the number of receives in the section. 
The algorithm generates the match pairs for the sends and receives in each section by a list of simple rules based on ranks \cite{DBLP:conf/kbse/HuangMM13}. A rank is a non-negative integer that represents the position of a send or a receive in a specific sequence in the section.

%Given the algorithm for match pair generation, the approach in this paper starts with $k=1$ and increments $k$ to a larger bound such that the number of the divided sections is changed, because the match pair generation for the new iteration is only meaningful if each process is sectioned differently. The growth of the size of the match pair set is polynomial as $k$ increases. The program property is then witnessed with the larger bound $k$; or the process repeats.

%contribution
To summarize, the main contributions of this paper include,
\begin{compactitem}
\item a new SMT encoding that is capable of proving the compatibility of synchronous send and detecting the violation of ready send,
\item an efficient algorithm that under-approximates the precise match pairs for a CTP, and
\item a set of benchmarks that demonstrate the new encoding with the under-approximation algorithm for match pairs is correct and is more efficient then a state-of-art MPI verifier. 
\end{compactitem}

The rest of the paper is organized as follows: 
Sections 2 presents two examples of CTP; Section 3 presents a few trivial rules for encoding MPI send modes; Section 4 discusses the new encoding and how it is able to prove synchronous send compatibility and to detect ready send violation;
Section 5 presents the general algorithm for match pair generation in the paper; Section 6 states that the precise match pairs can be generated; Section 7 gives the experimental results; Section 8 discusses the related work; and Section 9 is the conclusion and future work.


\begin{comment}
%!TEX root = paper.tex
\section{Introduction}
%Asynchronous message passing is a prevalent programming model in high performance computing (HPC). The model consists of two operations, send and receive, that are essential to message communication. The message communication is complicated because of the non-determinism that a receive may be matched with more than one send in the runtime. The use of match pair, a pair of a send and a receive that may potentially match in the runtime, is able to capture the message communication.
%The semantics can also be complicated because of two buffering settings in the runtime, infinite buffer (messages are buffered in the system) and zero buffer (no buffering in the system). Further, typical message passing standard such as message passing interface (MPI) specifies a special communication that uses collective operations to synchronize a program leading to a more complex program behavior. 
%Given the semantics, several common problems exist in message passing applications: the user-provided assertion may be violated in an execution; the program may deadlock for unexpected matching of receives; and the program may be incompatible with zero buffer semantics, meaning that no feasible schedule exists under zero buffer setting. 
Asynchronous message passing is a prevalent programming model in high performance computing (HPC).
The idea of message passing is simple at the outset; processes communicate by sending messages from one to another. It does not take long, however, to realize that despite the simplicity of the programming model, there is a lot of subtlety in message passing that can impact program behavior. For example, depending on the library, a programmer must be aware of such things as message non-determinism (concurrent sends to a process can arrive in any order), buffering semantics in the runtime (a process may block if the buffer is full), or broad synchronization operations that message with  groups of processes at the same time (i.e., collective operations). This inherent complexity means that the problems of determining if a message passing program is free of deadlock, compatible with a given buffering semantics, or if the message non-determinism affects the correctness of the computation are all NP-complete \cite{DBLP:conf/kbse/HuangMM13,HuangNFM15,DBLP:conf/fm/ForejtKNS14,deadlock-draft}. Showing any of these properties for an input program is hard.

Prior work on program correctness in the message passing paradigm can be roughly grouped into dynamic analysis where an existing runtime is manipulated to explore different scheduling outcomes \cite{DBLP:conf/ppopp/VakkalankaSGK08,DBLP:conf/sbmf/SharmaGB12}, model checking where a model of the original program is analyzed \cite{DBLP:conf/vmcai/Siegel07,DBLP:conf/pvm/Siegel07}, runtime verification where the program execution is observed but not manipulated \cite{DBLP:conf/sc/VetterS00,DBLP:conf/parco/KrammerBMR03,DBLP:conf/ptw/HilbrichSSM09}, and symbolic model checking where a model of the program is symbolically analyzed with an SMT solver \cite{DBLP:conf/kbse/HuangMM13,HuangNFM15}. This paper looks specifically at symbolic model checking and ways to better leverage the SMT solver in witnessing program properties for message passing programs.

The most efficient types of encoding for symbolic model checking of message passing programs, where efficient means they scale to programs with a high degree of message non-determinism, rely on the concept of a \emph{match pair} and a \emph{concurrent trace program} abstraction (CTP) \cite{DBLP:conf/kbse/HuangMM13,HuangNFM15,deadlock-draft,DBLP:conf/fm/ForejtKNS14,DBLP:conf/atva/ElwakilYW10,DBLP:conf/issta/ElwakilY10}. 
A match pair is a send and receive pair that may be matched by the run-time in some feasible execution of the program \cite{DBLP:conf/kbse/HuangMM13,HuangNFM15,deadlock-draft,DBLP:conf/fm/ForejtKNS14,DBLP:conf/atva/ElwakilYW10,DBLP:conf/issta/ElwakilY10}. 
A CTP is an abstraction of the program that removes all branching so each process is a sequence of send, receive, and wait (e.g., a witness to the completion of a send or receive) operations \cite{DBLP:conf/kbse/HuangMM13,HuangNFM15}. 
The abstraction is constructed by observing the execution of the program in the run time for a given input. Symbolic model checking gives the SMT solver an encoded CTP with a set of match pairs, and the solver then tries to select a subset of the given match pairs in such a way that the CTP deadlocks or violates an assertion. 

%The symbolic model checking requires a over-approximated set of match pairs as input. The set includes all the true match pairs and maybe a few match pairs that may never occur in the runtime. 
%The work then encodes the CTP into an SMT problem that uses a set of formulas to constrain the program behavior based on the semantics. The typical error such as deadlock, zero buffer incompatibility or assertion violation is then checked by solving the SMT problem. If a satisfying assignment exists for the SMT problem, then the error is detected where the match pairs are resolved to capture the message communication in a feasible program execution. If the SMT problem is unsatisfiable, the CTP is free of that error because no feasible schedule exists with any resolution of the match pairs. 

The size of the input match-pair set directly impacts the cost of the call to the SMT solver in the symbolic model checking of message passing programs; a large input set typically means an expensive solver call especially if there is a high-degree of message non-determinism. The idea in this paper is to reduce the cost of the solver call by reducing the size of the input match set to each call. Such a reduction cannot, however, be done naively; any match pair set must be \emph{message complete}.  A message complete match pair set for a CTP is one that lets the program run to completion (assuming it is free of deadlock). To reduce the number of match pairs and thereby decrease the cost of verification, this paper describes an iterative algorithm to successively generate under-approximations of the true set of match pairs until all match pairs are generated. The under-approximations retain the message complete property. 

%As such, reducing the size of the input match pair set directly impacts the run time cost, typically, of the SMT solver. A straightforward idea to improve the performance, therefore, is to reduce the number of the match pairs. A naive way of reducing the match pairs such as randomly selecting a subset of the match pairs, however, is not sufficient. The input set of match pairs must be sufficient to complete the message communication in a program execution if the CTP is free of deadlock. A match pair set that has this property for a given CTP is termed \emph{message complete}.
%The input match pairs must satisfy two properties: program completion (the program can run to completion assuming no deadlocks exist) and message non-determinism (the set has a certain degree of non-determinism such that a receive may be associated with more than one match pairs).  
%The algorithm in this paper explores a way to generate such a reduced set of match pairs.

To reduce the number of match pairs and thereby decrease the cost of verification, this paper describes an iterative algorithm to successively generate under-approximations of the true set of match pairs until all match pairs are generated. The under-approximations retain the message complete property. 
Such an approach focuses on property witnessing rather than proofs of correctness as the intent is to witness program properties early in the iterations before the set of match pairs gets large.
%Since the correctness of a CTP can not be proved given an under-approximated set of match pairs, this paper only discusses the capability of error detection for the SMT encoding. 
%The primary goal is to improve bug finding in a CTP; although, since the algorithm is iterative, it is able to eventually produce a set of match pairs that includes all reachable match pairs in the CTP, therefore it proves correctness. 


Generally, the algorithm generates the match pairs in two steps. 
First, each process is sectioned, where a section contains a sequence of receives in the process and a number of sends from all the senders that may potentially match these receives. The number of sends from each sender is roughly bounded by  $k$ given as a user input. 
%For each section, the algorithm precisely adds a set of sequences of sends from multiple potential sender processes. 
The sends from each sender are distributed sequentially. 
The total number of the sends distributed to the section is equal to the number of receives in the same section. 
%The number of the receives depends on three values that are statically known: the positive integer $k$ configured by the user, the number of all the potential sender processes and the number of the unmatched receives in the process. 
Match pairs are then generated for each section independently. This independent sectioning effectively ignores combinations of match pairs, that are feasible, but only available when the considered concurrent outstanding sends are more than the number of receives in the section. 
The algorithm generates the match pairs for the sends and receives in each section by a list of simple rules based on ranks \cite{DBLP:conf/kbse/HuangMM13}. A rank is a non-negative integer that represents the position of a send or a receive in a specific sequence in the section.

Given the algorithm for match pair generation, the approach in this paper starts with $k=1$ and increments $k$ to a larger bound such that the number of the divided sections is changed, because the match pair generation for the new iteration is only meaningful if each process is sectioned differently. The growth of the size of the match pair set is polynomial as $k$ increases. The program property is then witnessed with the larger bound $k$; or the process repeats.


%Have you added the discussion that k jumps by increments greater than 1? Meaning that you move k until the number of sections changes? In many regards, that is what is critical; start with k=1, then move k until the # of sections changes, run the solver, repeat. Is that discussion in the paper?


%The key insight of the solution is that each section is considered independently for match pair generation.  
%The algorithm does not generate the match pair for a send and a receive from two different sections, while in fact the send and the receive may match in the runtime. 
%As a result, the algorithm is able to under-approximate the precise match pairs as input to an SMT problem. 
%Due to the missed match pairs that are derived from matching sends and receives in different sections, the SMT problem may not encode the complete program behavior. However, if the problem is satisfiable, then it is known that a specific error exists for a feasible schedule that is resolved by the encoding. As such, the SMT encoding with a reduced set of match pairs can be used to detect errors. If the problem is, however, unsatisfiable for the under-approximated match pairs, it does not mean that there are provably no errors for any feasible execution in the runtime. It is then necessary to generate a larger set of match pairs by a new input K to capture more behavior for detection.


The paper includes the discussion that the precise match pairs for any CTP can be generated with a sufficiently large bound $k$. 
Experiments further show that the runtime cost of witnessing properties is dramatically reduced as the properties are witnessed with relatively small values of $k$ creating simple problem instances for the SMT solver. 
Experiments also show that the algorithm is able to scale to a program that employs a high degree of message non-determinism and/or a high degree of deep communication.
Deep communication in this context means that each sender for a receiver issues a long sequence of sends to the receiver.
The contributions include,
\begin{compactitem}
\item the efficient algorithm that under-approximates the precise match pairs for a CTP, 
%where the size depends on a positive integer $k$ configured by the user,
\item the proof that the precise match pairs for any CTP can be generated by the new algorithm with a sufficiently large bound $k$, and
\item the results from the experiments using a set of benchmarks to demonstrate the improved ability of witnessing program properties.
\end{compactitem}

The rest of the paper is organized as follows: 
Sections 2 presents the definition and semantics of CTP; Section 3 presents the general algorithm in the paper; Section 4 states that the precise match pairs can be generated; Section 5 gives the experimental results; Section 6 discusses the related work; and Section 7 is the conclusion and future work.

\end{comment}
